Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-06-03 talk/dls2014/paper/paper.tex
more explanations

In all cases, we remove the \lstinline!WRITE_BARRIER!  flag from the
object before we return. Thus, we never trigger the slow path again
before we do the next minor collection or we start the next
transaction (we always do a minor collection during a commit).

@@

Here, we will not go into detail about the integration of our STM
system with PyPy's JIT. In fact, we will disable it for all benchmarks
except those in section \ref{sec:performance-bench}. We would like to
regard it as a simple performance enhancement, but that is not what
happens in reality. First, since the JIT is a tracing
JIT\remi{explain?}  running in multiple threads, it may compile
different things in each run because of the non-deterministic
thread-scheduling of the operating system (OS). Second, it is able to
remove some allocations in some cases. Because compilation is already
non-deterministic, so is this allocation-removal. And third, we did
not have enough time to optimise integration with STM so that the JIT
exposes the overhead of STM more by speeding up all the rest.

@@

From the other end, there also have been attempts at lowering the
overhead of STM~\cite{warmhoff13,spear09} especially for a low number
of CPU cores. They have completely different approaches and it would
be interesting to see how well they work for our specific use
case. \remi{enough?}

>>>>

In all cases, we remove the \lstinline!WRITE_BARRIER!  flag from the
object before we return. Thus, we never trigger the slow path again
before we do the next minor collection or we start the next
transaction~\footnote{We always do a minor collection during a commit.}.

Note that we have three kinds of objects: \emph{young, overflow} and
\emph{old} objects. Young and overflow objects were created in the
same transaction; old objects always come from previous transactions.
The close integration between STM and the GC allows us to use this
information effectively: young objects do not trigger the slow path at
all, overflow objects only need to be handled by the GC part of the
write barrier once between collections, and only the old objects need
to be handled by the TM part once per transaction. This keeps the
overhead of the write barrier very low.

@@

Here, we will not go into detail about the integration of our STM
system with PyPy's JIT. In fact, we will disable it for all benchmarks
except those in section \ref{sec:performance-bench}. We would like to
regard it as a simple performance enhancement, but that is not what
happens in reality. First, since the JIT is a tracing JIT\cite{cfbolz09}
\remi{explain?}  running in multiple threads, it may compile
different things in each run because of the non-deterministic
thread-scheduling of the operating system (OS). Second, it is able to
remove some allocations in some cases. Because compilation is already
non-deterministic, so is this allocation-removal. And third, we did
not have enough time to optimise integration with STM so that the JIT
exposes the overhead of STM more by speeding up all the rest.

@@

From the other end, there also have been attempts at lowering the
overhead of STM~\cite{warmhoff13,spear09} especially for a low number
of CPU cores. \cite{spear09} works well for some workloads, in the
case of a dynamic language VM, we have to assume any possible
workload.  \cite{warmhoff13} uses a compile-time approach to select
from multiple optimised STMs. Their partitioned FastLane STM seems to
provide good performance already on two threads using master and
helper threads. It would be interesting to see if its design also
performs well in our use case.
